---
title: "Water Flow"
author: "Jared Flater"
date: "2/15/2021"
---


# Supp Figure S4

This plot shows the flow rate for each sample from each plot. 

If we plot MAB vs. flow, do prairie strips have an impact? 

```{r}
library(readxl)
library(tidyverse)
library(ggpubr)
library(rstatix)
```

```{r}
mabbarplot <- readRDS(file = "Data/mabbarplot.RDS")


flow <- read_excel("Data/WOR_Water_flow.xlsx")

flow$sample_number <- as.character(flow$sample_number)
```


```{r}
ggplot(flow, aes(x = as.factor(sample_number), y = `flow_g/s`, group = as.factor(plot))) +
  geom_line(position = position_dodge(width = .3)) +
  geom_point(aes(color = as.factor(treatment)), position = position_dodge(width = .3), size = 5) +
  geom_text(aes(label = plot), position = position_dodge(width = .3), show.legend = F) +
  theme_classic() +
  theme(
    legend.position = "top", 
    panel.background = element_rect(color = 'black', size = 1.4),
    strip.text.x = element_text(size = 10, face = 'bold'),
    strip.background = element_rect(colour = 'black', size = 1.4),
    axis.text.x = element_text(angle = 0, vjust = 0, hjust = 0.5)) +
    scale_color_viridis_d() +
    labs(x = "Time after runoff start (min)", y = "Flow (mL/s)", color = "Treatment \n(points labeled with plot)") +
    scale_x_discrete(labels = c("1" = "2.5", "2" = "7.5", "3" = "12.5", "4" = "17.5", "5" = "22.5", "6" = "27.5"))

ggsave(filename = "Figures/SupFig4.png", plot = last_plot(), device = "png", width = 6.5, height = 7.5, units = "in", dpi = 350)
```

# Supp Figure S5

```{r}
norm <- readRDS(file = "Data/norm.RDS") # Load norm data from ARGs.qmd

SixS <- norm %>%
  filter(Assay == "16S_Eub_338_518") 
SixS

SixS$treatment = factor(SixS$treatment, 
                             levels = c('WCS', 'WCSM', 'WCM'))
levels(SixS$treatment) <- list("CS"="WCS", "CSM"="WCSM", "CM"="WCM")
SixS$sample_number <- as.factor(SixS$sample_number)
levels(SixS$sample_number) <- list("2.5" = "1", "7.5" = "2", "12.5" = "3", "17.5" = "4", "22.5" = "5", "27.5" = "6")

my_comparisons <- list( c("WCS", "WCSM"), c("WCS", "WCM"), c("WCM", "WCSM"))

sig <- SixS %>%
  filter(!rep %in% c(4,5,6)) %>%
  droplevels()

sig
p <- ggplot(sig, aes(x=treatment, y=counts, fill = treatment)) + 
  geom_boxplot() +
  labs(x = "Treatment", y = "Log 10 gene 16s copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() +
  facet_wrap( ~ sample_number) +
  theme_pubr() +
  theme(
    legend.background = element_rect(
      fill = (alpha = 0),
      color = 'black',
      size = 0.1),
    panel.background = element_rect(color = 'black', size = 1.4),
    strip.text.x = element_text(size = 10, face = 'bold'),
    strip.background = element_rect(colour = 'black', size = 1.4),
    legend.position = "none") 
p  
```

```{r}
krusk.test <- sig %>%
  group_by(sample_number) %>%
  kruskal_test(counts ~ treatment)
krusk.test

stat.test <- sig %>%
  group_by(sample_number) %>%
  dunn_test(counts ~ treatment) %>%
  adjust_pvalue(method = "hochberg") %>%
  mutate(y.position = rep(c(log10(100000), log10(300000), log10(1000000)), 6))
stat.test
p <- ggboxplot(sig, x = "treatment", y = "counts", facet.by = "sample_number") +
  scale_y_continuous(trans = "log10")

options(scipen = 5)

p + 
  stat_pvalue_manual(stat.test, hide.ns = T) +
  theme_pubr() +
  labs(x = "Treatment", y = "Log 10 gene 16s copies",
       caption = get_pwc_label(stat.test)) +
  scale_x_discrete(labels = c("CSM" = "CSL", "CM" = "CL"))
ggsave(filename = "Figures/SuppFiggS5.png", plot = last_plot(), device = "png", units = "in", width = 6.5, height = 7.5, dpi = 300)
  
```

# Supp Figure S6

```{r}
df <- mabbarplot$data %>%
  group_by(plot, time) %>%
  mutate(SumAbund = sum(Abundance)) %>%
  rename(sample_number = time) %>%
  ungroup() %>%
  left_join(flow, by = c("plot", "sample_number")) %>%
  select(treatment.x, plot, sample_number, SumAbund, time_after_runoff_start, collection_time_sec, jare_tare_wt_g, jare_plus_runoff_wt_g, `flow_g/s`)

write_excel_csv(df, file = "Data/water_flow_with_MAB.csv")

lmplot <- df %>%
  filter(!treatment.x == "CS") %>%
  ggplot(aes(x = `flow_g/s`, y = SumAbund, color = treatment.x)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)  +
  scale_color_manual(values = c("darkgreen", "tan4"), name = "Treatment") +
  theme_bw()

lmplot
```

```{r}
# Two seperate lm's to calculated where they intersect
CM <- df %>%
  filter(!treatment.x %in% c("CS", "CSL")) %>%
  lm(SumAbund ~ `flow_g/s`, data = .)
summary(CM)$r.squared
summary(CM)$adj.r.squared

CSM <- df %>%
  filter(!treatment.x %in% c("CS", "CL")) %>%
  lm(SumAbund ~ `flow_g/s`, data = .)
summary(CSM)$r.squared
summary(CSM)$adj.r.squared

# where do the two lines intersect? 
cm <- rbind(coef(CM),coef(CSM)) # Coefficient matrix
c(-solve(cbind(cm[,2],-1)) %*% cm[,1])

lmplot <- df %>%
  filter(!treatment.x == "CS") %>%
  ggplot(aes(x = `flow_g/s`, y = SumAbund, color = treatment.x)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  geom_point(x = 7.373952, y = 0.286310, color = "black") +
  annotate("text", x = 8.1, y = 0.24, label = "7.37 (mL/s)") +
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")), 
    label.x = 3, show.legend = F) +
  labs(x = "Flow (mL/s)", y = "Relative abundance (%)") +
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) +
  annotate("segment", x = 7.373952, xend = 7.6, y = 0.286310, yend = 0.26,
  colour = "black") +
  scale_color_manual(name = "Treatment", labels = c("CSL", "CL"), values = c("darkgreen", "tan4")) +
  theme_bw() 

lmplot
ggsave(filename = "Figures/SupFig6.png", plot = last_plot(), device = "png", width = 6.5, height = 7.5, units = "in", dpi = 350)
```

# Supp Figure S7

If we plot MAB vs. flow, do prairie strips have an impact?

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(readr)

df <- read_csv("Data/water_flow_with_MAB.csv")

```

Look at differences in sample volume, time to collect
```{r}
colnames(df)

ggplot(df, aes(x = as.factor(sample_number), y = collection_time_sec, fill = as.factor(plot))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(~ treatment.x) +
  ggtitle("Sample Collection time")
ggplot(df, aes(x = as.factor(sample_number), y = jare_plus_runoff_wt_g - jare_tare_wt_g, fill = as.factor(plot))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(~ treatment.x) +
  ggtitle("Sample Volume")
```

A simple scatter plot looking at flow vs abundance, with a lm line fitted to each treatment.

```{r echo=FALSE}
# this data is at the end of the document
class(df$plot) 


plot <- df %>%
  filter(!treatment.x == "CS") %>%
  ggplot(aes(x = `flow_g/s`, y = SumAbund, color = treatment.x)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)  +
  scale_color_manual(values = c("darkgreen", "tan4"), name = "Treatment") +
  theme_bw()
plot
```

It appears that up to a flow of 7.5 g/s, strips may be effective at removing MAB in runoff water when compared to if no strips are present.

Calculate corrected (cor) MAB abundance by dividing MAB abundance by flow in each sample. Then calculate the treatment mean, sd, se and interquartile.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- df %>%
  filter(!treatment.x == "CS") %>%
  mutate(cor = SumAbund / `flow_g/s`) %>%# Make a plot
  group_by(treatment.x, time_after_runoff_start) %>%
  dplyr::summarise( 
    n = n(),
    mean = mean(cor),
    sd = sd(cor)
  ) %>%
  mutate(se = sd/sqrt(n))  %>%
  mutate(ic = se * qt((1-0.05)/2 + .5, n-1)) 
```

```{r echo=FALSE}
data %>%
  ggplot(aes(x = time_after_runoff_start, 
             y = mean, 
             ymin = mean - ic, 
             ymax = mean + ic,
             color = treatment.x,
             group = treatment.x)) +
  geom_point(position = position_dodge(width = 1), size = 2) +
  geom_line(position = position_dodge(width = 1), size = 2) +
  geom_errorbar(position = position_dodge(width = 1), width = .8) +
  labs(y = "Mean MAB / flow", x = "Time", title = "Flow weighted", subtitle = "Divide MAB relative abundance by flow") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("darkgreen", "tan4"), name = "Treatment") +
  theme_bw()
```

Here, despite the error bars, we have indications that when strips are present, there is less MAB in runoff water.

Event mean concentration (EMC)

EMC calculation

For a plot we have

1.  6 water samples

2.  Mass (vol) of water collected at each sample

3.  Time to collect each sample

4.  Count of ARGs filtered out of sample

The calculation

Event load (EL)


$$
=\frac{{T1ARGConc}*{T1Vol}+{T2ARGConc}*{T2Vol}+{T3ARGConc}*{T3Vol}+{T4ARGConc}*{T4Vol}+{T5ARGConc}*{T5Vol}+{T6ARGConc}*{T6Vol}}{Event TotalVol}
$$

```{r}
 # Load copy number from HT-qPCR runs
 # If you haven't calculated copy numbers yet,
 # see: \#\#\#\#\#\#\#
 water_counts <- read_csv("Data/WORWATERARGCOUNTS.csv")

 head(water_counts)
```